Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho, and Potential Impacts to Local Chinook Salmon Populations

### 2022-03-20 analysis/paper/paper.Rmd
Incorporated edits from reviewers IC and MR

We obtained an estimate of the number of juvenile Chinook Salmon consumed by dividing total biomass consumed by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG. The average weights of DSR and NRR emigrants were calculated from fish captured at seven rotary screw traps upstream of Deadwater Slough during the fall and spring periods. We also quantified the potential impact of Northern Pikeminnow predation on juvenile Chinook Salmon to adult returns by estimating the number of additional adult Chinook Salmon expected to return to Lower Granite Dam if predation by Pikeminnow was reduced or eliminated in the Deadwater Slough. We multiplied the estimated total juvenile Chinook Salmon consumed by the median Granite-to-Granite smolt-to-adult return rate (SAR) of 0.00614 (SD = 0.00051) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.

@@

Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` in fall 2019 and from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` in fall 2020 (Table \@ref(tab:abund-tab)). For both surveys, the unadjusted multiple-census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)). Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, we deemed the Schnabel estimates most appropriate. Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`). Using the delayed-mixing Schnabel estimator, the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`). For spring 2021, we estimated `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough. Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm and areal densities (fish/100 m$^2$) of `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.

>>>>

The estimate of total juvenile Chinook Salmon biomass consumed was then converted to the total estimated number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG. The average weights of DSR and NRR emigrants were calculated from fish captured at seven rotary screw traps upstream of Deadwater Slough during the fall and spring periods. Although the primary impact to Chinook Salmon in Deadwater Slough by Northern Pikeminnow is juvenile predation, the common metric to evaluate salmon recovery is often adult returns. We assessed the potential impacts on adult returns by estimating the number of additional adult Chinook Salmon expected to return to Lower Granite Dam if predation was reduced or eliminated in the Deadwater Slough. This calculation was made by multiplying the total estimated juvenile Chinook Salmon consumed by the median Granite-to-Granite SAR of 0.00614 (sd = 0.00051) from @McCann2019 for Upper Salmon River MPG Chinook Salmon.

@@

Mark-recapture abundance estimates of Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` depending on the estimator used (Table \@ref(tab:abund-tab)). For both surveys, the unadjusted multiple-census estimates were larger than the single census estimates and small numbers of recaptures resulted in wide confidence intervals (Figure \@ref(fig:abund-fig)). Because our sampling design most closely matched a multiple census estimator and required a small sample size modification, we deemed the Schnabel estimates most appropriate. Accordingly, the mean Northern Pikeminnow abundance for the two fall sampling events was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_F, spring_abund$Uci_F)), big.mark = ","), collapse = " - ")`). Using the delayed-mixing Schnabel estimator, the mean fall abundance estimate was `r prettyNum(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` (95% CI: `r prettyNum(mean(all_est$Lci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")` - `r prettyNum(mean(all_est$Uci[all_est$model == "Schnabel - Delayed Mixing"]), big.mark = ",")`). For spring 2021, we estimated `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough. Those estimates translate to linear densities of `r prettyNum(round(mean(all_est$N[all_est$model == "Schnabel - Delayed Mixing"]) / ddw_rkms), big.mark = ",")` and `r prettyNum(round(spring_abund$N_S / ddw_rkms), big.mark = ",")` Northern Pikeminnow per rkm and areal densities (fish/100 m$^2$) of `r round((mean(all_est$N[all_est$model == "Schnabel"]) / ddw_m2) * 100, 1)` and `r round((spring_abund$N_S / ddw_m2) * 100, 1)` for fall and spring, respectively.
